---
title: "Current status check"
author: "Ivo Borkus"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
    number_sections: yes
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = T, message = FALSE, warning = FALSE,
  fig.align = "center", fig.width = 10, fig.height = 10
)
knitr::render_markdown()
```

```{css, echo = F}
pre code, pre, code {
  white-space: pre !important;
  overflow-x: scroll !important;
  word-break: keep-all !important;
  word-wrap: initial !important;
}

pre {
  max-height: 300px;
  overflow-y: auto;
}
```
```{r loading libraries, include = FALSE}
library(Seurat)
library(dplyr)
library(ggplot2)
library(here)
library(tidyr)
library(patchwork)
library(UCell)
library(qs2)
library(future)
plan(multisession, workers = 7)
options(future.globals.maxSize = 100000 * 1024^2)
# source(here("Brain-Met_V2/functions.R"))
source(here('Glioma_Project/Scripting/utils.R'))
source(here('Glioma_Project/Scripting/functions.R'))
# source(here("External_data/Fastq_processing/metadata_functions.R"))
```

# Setting paths and names

Here we set the variables necessary to run my custom scripts. The rest of the script therefore should not be tweaked. 

However, as there is an interpretation stage, all code after Results is custom for each object.
```{r}
exc_path <- here("Glioma_Project/01_data/Excel")
rmd_path <- here("Glioma_Project/02_knits/figs")
obj_path <- here("Glioma_Project/01_data")
fig_path <- here("Glioma_Project/04_figs")

object_name <- "Processed"
reduction_name <- "Inhouse_Harmony"
Processed_name <- "Harmony"
sample_col <- "sample_name"
batch_cols <- c(
  "tumour_full", "tumour_abbreviated", "grade", "Cell_Type", "Cell_Subtype", "sample_name", "IDH_status",
  "FACS_sorting", "tier_correct", "cell_het_paper_id", "storage_condition", "prior_treatment", "gender", "Tumour_Margin",
  "Primary_or_Recurrent", "Previous_Surgery", "Radiotherapy", "Chemotherapy", "MGMT_Methylation"
)
continous_cols <- c("nCount_RNA", "nFeature_RNA", "percent.mt", "age_at_surgery_correct", "nuclear_fraction")


# subsetting <- c("Cell_Type", c("Myeloid", "Dendritic cells", "Mast cells")) #  Subset-col, cells_of_interest
subsetting <- c(NULL, NULL) #  Subset-col, cells_of_interest

norm_meth_scale <- c(TRUE, "LogNormalize", 10000) # Perform log normalisatin, method, scale
vf <- c("vst", 2000, FALSE, FALSE) # VF method, amount, use old vf method, silence VDJgenes using scRep
scaling <- TRUE # w
var.to.regress <- NULL # supply the var to be regressed if valid
harmony <- T # Run harmony or not
n_PCS <- 50 # Amount of Pcs to include
clustering <- c(TRUE, 4) # Perform clustering, Algorithm
sub_dir <- NULL


feature_vec_chosen <- feature_vec_overall
```

```{r}
seurat_obj_path <- ifelse(!is.null(sub_dir), here(obj_path, sub_dir, paste0(Processed_name, ".qs2")), here(obj_path, paste0(Processed_name, ".qs2")))
seurat_obj <- qs_read(seurat_obj_path)
.size(seurat_obj)
```

```{r, eval = F}
meta_df <- seurat_obj@meta.data
meta_df[,c(9:30,38:40)] %>% distinct() -> new_df
new_df %>% tibble::remove_rownames() -> distinct_meta
check <- lapply(colnames(distinct_meta),\(c){if(class(distinct_meta[[c]]) == 'list'){distinct_meta[[c]] <- as.character(distinct_meta[[c]])} else {distinct_meta[[c]]} }) %>% as.data.frame()
colnames(check) <- colnames(distinct_meta)
write.csv(check, here('Glioma_Project/01_data/00_metadata.csv'))

seurat_obj@meta.data$file_name %>% unique()
(seurat_obj$cell_het_paper_id == 'ODG-7') %>% table() # Strange
```

```{r cleaning the objects}
p <- VlnPlot(seurat_obj, features = c('nCount_RNA','nFeature_RNA','percent_mt'), group.by = 'Ann_lvl_2', ncol = 2) 
# pdf(here('Glioma_Project/04_figs/00Test.pdf'), width= 15, height = 20)
# p
# dev.off()
Idents(seurat_obj) <- 'Ann_lvl_2'
Myeloid_anns <- c("Mg_classic", "Mg_Inflam", "Mg_Anti-Inflam", "Mg_PLCG2", "Mg_active", "BAMs", "TAMs", "Ma-Inflam", "Ma-IFN", "Ma-Mt+", "Mono-Classic","Myeloid_Prolif","DCs_CD1C+","pDCs",'Mast-cells')
T_anns <- c('CD8+_TNFSF9','CD8+_Cytotoxic','CD8+ Eff-Exh','CD4+ cm','CD4+ Tfh','Tregs','T-IFN','T_Prolif','MAIT-cells','NK')
seurat_obj_M <- subset(x = seurat_obj, idents = Myeloid_anns, invert = F)
seurat_obj_T <- subset(x = seurat_obj, idents = T_anns, invert =F)

obj <- seurat_obj_T
reduction_name <- 'T'
samples_to_use <- obj@meta.data %>% group_by(sample_name) %>% summarise(count = n()) %>% filter(count> 1) %>% select(sample_name) %>%unlist(as.vector(.))# as.vector()
obj <- subset(obj, subset = sample_name %in% samples_to_use)
obj[["RNA"]] <- split(obj[["RNA"]], f = obj@meta.data[["sample_name"]])
obj <- FindVariableFeatures(obj, selection.method = 'vst', nfeatures = 1000, verbose = F)
VariableFeatures(obj) <- .remove_VDJ(VariableFeatures(obj))
obj <- ScaleData(object = obj, features = VariableFeatures(obj), verbose = FALSE)
obj <- RunPCA(
        object = obj, features = VariableFeatures(obj),
        nfeatures.print = 5, ndims.print = 1:2,
        reduction.name = paste0("pca_", reduction_name),
        npcs = 100
    )
# p1 <-.seurat_dimplot(obj, reduction = 'pca_T',label = T, title = 'Annotation T cells',group = 'Ann_lvl_2', pt.size = 1) + theme(legend.position = 'none')
# p1 <-DimPlot(obj, reduction = 'pca_T',dims = c(3,4),label = T,group = 'Ann_lvl_2', pt.size = 1) + theme(legend.position = 'none')
ElbowPlot(obj, reduction = paste0("pca_", reduction_name), ndims = 100)
VizDimLoadings(obj, dims = 1:5, reduction = paste0("pca_", reduction_name), nfeatures = 15)
n_PCS <- 20
obj <- RunPCA(
        object = obj, features = VariableFeatures(obj),
        nfeatures.print = 5, ndims.print = 1:2,
        reduction.name = paste0("pca_", reduction_name),
        npcs = n_PCS
    )

obj <- IntegrateLayers(
            object = obj, method = HarmonyIntegration,
            orig.reduction = paste0("pca_", reduction_name), new.reduction = paste0("harmony_pca.", reduction_name),
            verbose = TRUE
        )
obj <- FindNeighbors(obj, reduction = paste0("harmony_pca.", reduction_name), dims = 1:as.numeric(n_PCS))
res_values <- c(0.01, 0.05, 0.1, 0.2, 0.3, 0.5, 0.7, 0.9, 1.5)
obj <- FindClusters(obj, resolution = res_values, algorithm = clustering[2])
obj <- RunUMAP(obj, reduction =  paste0("harmony_pca.", reduction_name), dims = 1:as.numeric(n_PCS), reduction.name =  paste0("harmony_umap.", reduction_name))
obj <- JoinLayers(obj)



p1 <-.seurat_dimplot(obj, reduction = paste0("harmony_umap.", reduction_name),label = T, title = 'Annotation T cells',group = 'Ann_lvl_2', pt.size = 1) #+ theme(legend.position = 'none')
resolution_columns <- grep("^RNA_snn_res\\.",
    colnames(obj@meta.data),
    value = TRUE
)
p2 <- .seurat_dimplot(obj, reduction = paste0("harmony_umap.", reduction_name), group = resolution_columns, label = TRUE) & NoLegend()
p1 + p2
# p1 <-.seurat_dimplot(obj, reduction = paste0("harmony_umap.", reduction_name),label = T, title = 'Annotation T cells',group = 'sample_name', pt.size = 1) + theme(legend.position = 'none')


```

# To Do
RE: pro/anti-inf signatures, I think it would make sense to compute the overall 
signatures only on myeloid cells to compare the three diseases but you could also stratify then into
 microglia, macrophages and monocytes (DCs could be included but I don't think this signatures really apply to them).
Then if you want to add significance you could either average the scores by patient (only myeloid) 
or pseudobulk only myeloid per patient and score at a pseudobulk level (I would prefer this approach), 
then you can apply a max rank sum test on the score per patient between groups (disease/IDH-status...).

```{r pro anti sign}
immune_markers
gene_list <- list(
    inflammatory = c("CCL4", "IL1B", "CCL3", "RELB", "ICAM1", "TNFAIP3"), #  Juan microglia paper
    Anti_inflammatory = c("CD163", "SELENOP", "MRC1") # Juan microglia paper
)

# old <- list(
#   BD = c("LYZ", "EREG", "S100A8", "S100A9", "ITGAL", "ITGA4", "PTPRC", "FCGR3A", "CEBPD", "EGR2", "IRF4", "RUNX1", "STAT3", "CCL22", "CXCL2", "CD14", "S100A4", "S100A6", "FCN1", "VCAN", "CD36", "LGALS3", "MRC1", "CD163", "SELENOP", "SLC40A1", "CXCR4"),
#   MG = c("P2RY12", "CX3CR1", "CSF1R", "AXL", "TMEM119", "CD84", "CSF3R", "ELMO1", "MEF2A", "BIN1", "SCIN", "FCGBP", "OLFML3", "NAV3", "SALL1", "SIGLEC8", "SLC1A3", "ITGAM", "TREM2", "PROS1", "SPARC", "PLCL1", "VIM", "APOE", "SMAD3", "P2RX7", "LYVE1", "MS4A7", "MRC1", "FOLR2", "CD163", "HLA-DRA", "TIMD4", "ADGRG1", "TGFBI", "CD74")
  # Mast = c("IL1RL1", "HDC", "TNFSF10", "FGL2", "IL17RB", "CCL2", "CSF1", "IL13"),
  # pDCs = c("GZMB", "IGJ", "SERPINF1", "ITM2C", "IRF7", "TCF4", "BCL11A", "LILRA4", "MZB1", "DERL3", "IL3RA", "ZFAT", "NRP1", "CLEC4C", "IRF8", "SLC15A4", "BLNK"),
  # DC_general = c("CD1C", "CLEC9A"),
  # Prolif = c("UBE2C", "MKI67", "TOP2A"),
  # Mg_homeostatic = c("PTPRC", "ITGAM", "P2RY12"),
  # Mg_activated = c("CX3CR1", "GPR34", "FOXP2"),
  # Mg_Reslike = c("PLCL1", "CEBPD", "VIM"),
  # Mg_stress = c("HSPA1A", "HSPA1B", "SORCS2"),
  # Mg_Phag = c("GPNMB", "RGCC", "PPARG"),
  # Mg_Inflam = c("CCL4", "IL1B", "CCL3"),
  # Mg_Inflam_ICAM1 = c("ICAM1", "RELB", "TNFAIP3"),
  # Mg_INFg = c("IFNG", "IFIT2", "IFIT3", "STAT1"),
  # BAM = c("F13A1", "LYVE1", "CD36"),
  # BMD_Anti_Inflam = c("CD163", "SELENOP", "MRC1")
# )
```

```{r residency sig}
BD <-  c("LYZ", "EREG", "S100A8", "S100A9", "ITGAL", "ITGA4", "PTPRC", "FCGR3A", "CEBPD", "EGR2", "IRF4", "RUNX1", "STAT3", "CCL22", "CXCL2", "CD14", "S100A4", "S100A6", "FCN1", "VCAN", "CD36", "LGALS3", "MRC1", "CD163", "SELENOP", "SLC40A1", "CXCR4")
MG <- c("P2RY12", "CX3CR1", "CSF1R", "AXL", "TMEM119", "CD84", "CSF3R", "ELMO1", "MEF2A", "BIN1", "SCIN", "FCGBP", "OLFML3", "NAV3", "SALL1", "SIGLEC8", "SLC1A3", "ITGAM", "TREM2", "PROS1", "SPARC", "PLCL1", "VIM", "APOE", "SMAD3", "P2RX7", "LYVE1", "MS4A7", "MRC1", "FOLR2", "CD163", "HLA-DRA", "TIMD4", "ADGRG1", "TGFBI", "CD74")
```


```{r computing signatures on subsets}

seurat_obj <- AddModuleScore_UCell(seurat_obj,
    features = gene_list
)

```

RE: Milo, would make more sense to split Lymphoid and Myeloid cells and maybe,
just for the plots, split the myeloids into microglia, macrophages, monocytes and DCs
and the lymphoid into CD4, CD8, NK and B cells if any? 
and I would definitely remove the bad quality and the Hb+ cells as they will only add more noise (from all of the analyses actually).
Plus let's try to figure out what this mixed population is (doublets? proliferative?),
otherwise I would also remove them as it is not providing any helpful information. 
You can also run scCODA to validate these findings both for the diseases and the IDH mut/wt.

